---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading required libraries
```{r echo=TRUE}
library(ggplot2)
```

## Loading and preprocessing the data
```{r echo=TRUE}
# Load the data
unzip(zipfile = "./activity.zip", exdir = "./data")
rawData <- read.csv(file="./data/activity.csv",
                    colClasses = c("numeric", "Date", "numeric"))
data <- rawData[complete.cases(rawData), ]
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
# Calculate the total number of steps taken each day
# (removed the missing value) 
totalSteps <- aggregate(data$steps, by=list(data$date), FUN=sum)
colnames(totalSteps) <- c("date", "total")

hist(x=totalSteps$total, main="Histogram of the total number of steps taken each day", xlab="Total number of steps")
```

```{r echo=TRUE}
# Calculate and report the mean and median total number of steps taken per day
meanTotalSteps <- mean(totalSteps$total)
medianTotalSteps <- median(totalSteps$total)

# Convert to string
strMeanTotalSteps <- paste(meanTotalSteps)
strMedianTotalSteps <- paste(medianTotalSteps)
```

Mean total number of steps taken per day is `r strMeanTotalSteps`.  
Median total number of steps taken per day is `r strMedianTotalSteps`.  

## What is the average daily activity pattern?
```{r echo=TRUE}
avgStepsInterval <- aggregate(data$steps, by=list(data$interval), 
                              FUN=mean)
colnames(avgStepsInterval) <- c("interval", "average")

g <- ggplot(avgStepsInterval, aes(x = interval, y = average)) 
g <- g + geom_line()
g <- g + labs(x = "Interval")
g <- g + labs(y = "Average number of steps")
g <- g + labs(title = "Interval-based average number of steps")
print(g)
```

```{r echo=TRUE}
maxAvgStepsInterval <- max(avgStepsInterval$average)
interval <- avgStepsInterval[avgStepsInterval$average == maxAvgStepsInterval, 1]
```

Interval `r interval` contains the maximum of average number of steps.  

## Imputing missing values
```{r echo=TRUE}
# Calculate the total number of missing values in the dataset
totalNAs <- sum(!complete.cases(rawData))
```

The total number of rows that have missing values is `r totalNAs`.  

```{r echo=TRUE}
# To fill in all missing values, using the mean for these intervals should help reduce bias and can represent the whole population.
processedData <- rawData
for(index in 1:nrow(processedData)) {
    if(is.na(processedData[index, 1])) {
        interval <- processedData[index, 3]
        avgStep <- avgStepsInterval[avgStepsInterval$interval == interval, 2]
        processedData[index, 1] <- avgStep
    }
}

# Calculate the total number of steps taken each day on processed data
totalProcessedTotalSteps <- aggregate(processedData$steps, 
                                      by=list(processedData$date), 
                                      FUN=sum)
colnames(totalProcessedTotalSteps) <- c("date", "total")

hist(x=totalProcessedTotalSteps$total, 
     main="Histogram of the total number of steps on processed data", 
     xlab="Total number of steps")
```

```{r echo=TRUE}
# Calculate and report the mean and median total number of steps on processed data
meanProcessedTotalSteps <- mean(totalProcessedTotalSteps$total)
medianProcessedTotalSteps <- median(totalProcessedTotalSteps$total)

# Convert to string
strMeanProcessedTotalSteps <- paste(meanProcessedTotalSteps)
strMedianProcessedTotalSteps <- paste(medianProcessedTotalSteps)
```

Mean total number of steps on processed data is `r strMeanProcessedTotalSteps`.  
Median total number of steps on processed data is `r strMedianProcessedTotalSteps`.  

Compared histograms of the data without NAs and filling NAs with the mean of steps in each interval, the data distribution is simialr. The median of the data filling NAs with the mean of steps in each interval is the same as the mean of the data without NAs. In conclusion, the impact of imputing missing data is not significant.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
# Group data into two groups: weekday and weekend
processedData$day <- "weekday"
for(index in 1:nrow(processedData)) {
    day <- weekdays(processedData[index, 2])
    if(day == "Saturday" | day == "Sunday") {
        processedData[index, 4] <- "weekend"
    }
}
processedData[, 'day'] <- as.factor(processedData[, 'day'])

# Calculate the average number of steps grouped by weekday/weekend and 5-minute intervals.
avgTotalStepsDay <- tapply(processedData$steps, 
                           list(processedData$interval, processedData$day), 
                           mean)
avgTotalStepsDay <- as.data.frame(as.table(avgTotalStepsDay))
colnames(avgTotalStepsDay) <- c("interval", "day", "average")

g <- ggplot(avgTotalStepsDay, aes(x = interval, y = average))
g <- g + geom_line(aes(group=1)) # Group all points; otherwise no line will show
g <- g + facet_grid(day ~ .)
g <- g + labs(x = "Interval")
g <- g + labs(y = "Average number of steps")
g <- g + labs(title = "Interval-based average number of steps grouped by day")
print(g)
```
