---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading required libraries
```{r echo=TRUE}
library(ggplot2)
```

## Loading and preprocessing the data
```{r echo=TRUE}
# Load the data
unzip(zipfile = "./activity.zip", exdir = "./data")
rawData <- read.csv(file="./data/activity.csv",
                    colClasses = c("numeric", "Date", "numeric"))
data <- rawData[complete.cases(rawData), ]
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
# Calculate the total number of steps taken each day
# (removed the missing value) 
totalSteps <- aggregate(data$steps, by=list(data$date), FUN=sum)
colnames(totalSteps) <- c("date", "total")

hist(x=totalSteps$total, main="Histogram of the total number of steps taken each day", xlab="Total number of steps")
```

```{r echo=TRUE}
# Calculate and report the mean and median total number of steps taken per day
meanTotalSteps <- mean(totalSteps$total)
medianTotalSteps <- median(totalSteps$total)

# Convert to string
strMeanTotalSteps <- paste(meanTotalSteps)
strMedianTotalSteps <- paste(medianTotalSteps)
```

Mean total number of steps taken per day is `r strMeanTotalSteps`.  
Median total number of steps taken per day is `r strMedianTotalSteps`.  

## What is the average daily activity pattern?
```{r echo=TRUE}
avgStepsInterval <- aggregate(data$steps, by=list(data$interval), 
                              FUN=mean)
colnames(avgStepsInterval) <- c("interval", "average")

g <- ggplot(avgStepsInterval, aes(x = interval, y = average)) 
g <- g + geom_line()
g <- g + labs(x = "Interval")
g <- g + labs(y = "Average number of steps")
g <- g + labs(title = "Interval-based average number of steps")
print(g)
```

```{r echo=TRUE}
maxAvgStepsInterval <- max(avgStepsInterval$average)
interval <- avgStepsInterval[avgStepsInterval$average == maxAvgStepsInterval, 1]
```

Interval `r interval` contains the maximum of average number of steps.  

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
